---
title: "Battle of the Bands: Text Mining Lyrics from Rush, Queen, and AC/DC"
subtitle: "Section 2: Data Exploration"
author: "Michael Foley"
date: "`r Sys.Date()`"
output: 
  html_document:
    css: "style.css"
    theme: flatly
    toc: true
    toc_float: true
    highlight: haddock
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidytext)
library(plotly)
library(glue)
library(janitor)

lyrics <- readRDS("./1_lyrics.Rds")
lyrics_lines <- readRDS("./1_lyrics_lines.Rds")

band_palette <- c("Queen" = "lightgoldenrod", "Rush" = "darkseagreen", "AC/DC" = "slategray1")
```

> Section 2 of my [Battle of the Bands](https://mpfoley73.github.io/battle-of-the-bands/) text mining project is a brief exploration of the lyrics dataset to get acquainted with the data and perform any transformations that might be useful in subsequent analyses. 

The dataset consists of `r nrow(lyrics)` songs spread fairly even across the three bands.

```{r}
lyrics %>% 
  group_by(band) %>%
  summarize(
    albums = n_distinct(album),
    songs = n()
  ) %>%
  adorn_totals() %>%
  adorn_percentages(denominator = "col") %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  flextable::flextable() %>%
  flextable::autofit()
```

## Song Counts

The bands have substantial overlap in productive years. All three bands had a flurry of output over their first half-decade or so, then gradually slowed their pace. AC/DC is still churning out music!

```{r}
lyrics %>% 
  count(band, released, album) %>%
  mutate(
    released = case_when(
      album == "Fly by Night" ~ as.integer(1974), 
      album == "Queen" ~ as.integer(1972),
      album == "Queen II" ~ as.integer(1973),
      TRUE ~ released)) %>% 
  ggplot(aes(x = released, y = n, fill = band)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = band_palette) +
  geom_text(aes(label = album, y = 0.5), angle = 90, hjust = "bottom", vjust = .25,
            size = 3, color = "grey50") +
  scale_x_continuous(limits = c(1970, 2025), 
                     breaks = seq(1970, 2025, by = 5), 
                     minor_breaks = 1970:2025, 
                     expand = c(0,0)) +
  theme_light() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  ) +
  labs(x = "Released", y = "Songs on Album", 
       title = glue("{n_distinct(lyrics$album)} albums, {nrow(lyrics)} songs.")) +
  facet_grid(rows = vars(band))
```

# Line Counts

```{r}
lyrics_1 <- lyrics %>%
  inner_join(lyrics_lines %>% count(song_id, name = "n_lines"), by = "song_id")

lc_quant_by_band <- lyrics_1 %>%
  split(lyrics_1$band) %>% 
  lapply(function(x) quantile(x$n_lines)) %>%
  bind_rows(.id = "band")

lc_quant <- quantile(lyrics_1$n_lines)
lc_quant_df <- lc_quant %>% as_tibble_row() %>% bind_cols(band = "Overall")
lc_quant_tbl <- bind_rows(lc_quant_by_band, lc_quant_df)

shortest_songs <- lyrics_1 %>% slice_min(order_by = n_lines) %>% pull(song)
longest_songs <- lyrics_1 %>% slice_max(order_by = n_lines) %>% pull(song)
```

The overall IQR of lines per song was `r lc_quant["25%"]` to `r lc_quant["75%"]` lines, and those quantiles were consistent across the bands. 

```{r}
lc_quant_tbl %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_caption("Song Lines Quantiles")
```

Queen wrote the shortest songs. `r shortest_songs[1]` and `r shortest_songs[2]` where `r min(lyrics_1$n_lines)` lines. The longest song was `r longest_songs` by AC/DC (`r max(lyrics_1$n_lines)` lines).

```{r}
p <- lyrics_1 %>% 
  ggplot(aes(x = released, y = n_lines, group = as.factor(released), color = band,
             text = glue("Band: {band} <br>",
                         "Album: {album} <br>",
                         "Lyrics: {writer} <br>",
                         "Song: {song} <br>",
                         "Lines: {n_lines}"))) +
  geom_boxplot() +
  geom_jitter(height = 0, size = 2, alpha = 0.6) +
  # scale_fill_manual(values = band_palette) +
  scale_color_manual(values = band_palette) +
  theme_light() +
  theme(legend.position = "none") +
  labs(
    x = NULL, y = "Line Count",
    title = glue("Line count IQR is {lc_quant['25%']} to {lc_quant['75%']} lines.")
  ) +
  facet_wrap(vars(band))

ggplotly(p, tooltip = "text")
```

# Word Counts

The **tidytext** package splits the text into "tokens" (words). 

```{r}
word_count <- lyrics_1 %>% 
  unnest_tokens(output = "word", input = "lyrics", token = "words") %>%
  count(song_id, name = "n_words")

lyrics_2 <- lyrics_1 %>%
  inner_join(word_count, by = "song_id")

wc_quant_by_band <- lyrics_2 %>%
  split(lyrics_1$band) %>% 
  lapply(function(x) quantile(x$n_words)) %>%
  bind_rows(.id = "band")

wc_quant <- quantile(lyrics_2$n_words)
wc_quant_df <- wc_quant %>% as_tibble_row() %>% bind_cols(band = "Overall")
wc_quant_tbl <- bind_rows(wc_quant_by_band, wc_quant_df)

shortest_wc_song <- lyrics_2 %>% slice_min(order_by = n_words) %>% pull(song)
longest_wc_song <- lyrics_2 %>% slice_max(order_by = n_words) %>% pull(song)
```

The overall IQR of words per song was `r wc_quant["25%"]` to `r wc_quant["75%"]` lines. AC/DC and Queen had similar quantiles, and Rush songs were about 10% longer.

```{r}
wc_quant_tbl %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_caption("Song Words Quantiles")
```

The IQR of words per song was `r wc_quant["25%"]` to `r wc_quant["75%"]` words. The shortest song was `r shortest_wc_song` (`r min(lyrics_2$n_words)` words). The longest song was `r longest_wc_song` at (`r max(lyrics_2$n_words)` words).

```{r}
p <- lyrics_2 %>% 
  ggplot(aes(x = released, y = n_words, group = as.factor(released), color = band,
             text = glue("Band: {band} <br>",
                         "Album: {album} <br>",
                         "Lyrics: {writer} <br>",
                         "Song: {song} <br>",
                         "Lines: {n_lines} <br>",
                         "Words: {n_words}"))) +
  geom_boxplot() +
  geom_jitter(height = 0, size = 2, alpha = 0.6) +
  scale_color_manual(values = band_palette) +
  theme_light() +
  theme(legend.position = "none") +
  labs(
    x = NULL, y = "Word Count",
    title = glue("Word count IQR is {wc_quant['25%']} to {wc_quant['75%']} words.")
  ) +
  facet_wrap(vars(band))

ggplotly(p, tooltip = "text")
```

## Save Work

Save the lyrics with summary stats for subsequent steps.

```{r}
saveRDS(lyrics_2, "./2_lyrics.Rds")
```
